---
title: "Comparing Dice Rolls"
author: "Norm Bryar"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{DnD Dice}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
The DnD Dice Package has some hand utility functions to help you, 
as a player, examine the distribution of various dice-roll combinations.  This
helps you decide which feats, weapons, fighting-styles, etc. to pick
to match your play-style and expected enemy-characteristics.  

It also has some helper-functions for you as a D.M. to create challenging
encounters for a party of various levels.

# Creating Die Roll Combinations
Die rolls are modeled here with  
- a constant,  
- list of dice-counts  
- list of dice-types  
e.g. `rollDice( dieCt=c(1,1), dieType=c(8,4), adj=1 )`

Typically, you build-up higher-order functions via the `makeRoll` sibling-function:  

```{r, results='as-is'}
library(dungeonDice)
poisonLongSword <- makeRoll( c(1,1), c(8,4), 1 )  
poisonLongSword()
```  


## Composing Die Types
The simplest composition involves the `multiRoll` helper, e.g.  
`rapierDag2wf <- multiRoll( c(makeRoll( 1, 8, 2 ), makeRoll(1, 4)) )  
rapierDag2wf()`

This adds a 1D8+2 rapier damage with a bonus-action attack of 1D4 dmg.  (Assuming, for now, both hit).

### Composing Policies

An optional eval function can model other compositions besides sum, e.g. 

`advantageTohit <- multiRoll( 
                       c( multiRoll( c(makeRoll(1,20), makeRoll(1,20)), max ),
                          constRoll( 3 ) ) )`  
                          
this takes the better of two 1D20 rolls and adds your +3 proficiency bonus.

Other helper-utilities let you compose conditional results, e.g. for a Paladin's 
Great Weapon Fighting style-option, you can model a greatsword's damage as 

`oneTwoRerollD6 <- ifSuccessRoll( function(r) {r>2}, makeRoll(1,6), makeRoll(1,6) )
 greatSwordGWF  <- multiRoll( c(oneTwoRerollD6,oneTwoRerollD6), sum )`


## Common Examples
Attribute Rolls (best 3 of 4D6): `sum( sort( genSamples( 4, makeRoll( 1,6,0 ) ), decreasing=TRUE )[1:3] )` (which we've exported as attribRoll())

Advantage Roll (best 1 of 2D20): `multiRoll( c(rollFunc, rollFunc), max )` (a.k.a. export advantageRoll)

Reliable Taleent (<10 = 10): `multiRoll( c(makeRoll(1,20,0), function(){10}), max )` (a.k.a. reliableTalentRoll)

## Viewing Die Roll Distributions
What do you suppose this looks like? 
```{r, fig.show='hold', fig.cap='important plots'}
plot(1:10)
plot(10:1)
```

## Simulations
Of course, raw damage is only part of the story.  You may want to decide whether your 4th-level Rogue
should take a DEX gain for +1 ToHit and + 1 Dmg vs. the Savage Attacker feat's advantage to damage.  Probability to hit needs figure into the distribution (as does DEX's affect on Armor Class, Sneak,
DEX saving-throws, etc., but lets keep this simulation simple for now).

`rapierDex16  <- makeRoll( 1, 8, (2+3) )  # proficiency 2, dex mod 3
 hitAC16Dex16 <- toHit( 16, 2, 3 )
 dex16AtkDmg  <- ifelseRoll( hitAC16Dex16, rapierDex16, constRoll(0) )

 rapierDice   <- makeRoll( 1, 8,   0   )  # Savage Attacker re-rolls die only
 savageRapier <- multiRoll( c(advantageRoll( rapierDice ),
                              constRoll(2 + 2) ), sum )
 hitAc16Dex14 <- toHit( 16, 2, 2 )
 savageAtkDmg <- ifelseRoll( hitAc16Dex14, savageRapier, constRoll(0) )
 
 sim <- abSamples( 10000, dex16AtkDmg, savageAtkDmg )
 showExp( sim, "Rapier: DEX 16 v. DEx 14 savage" )`
 
 The red 'control' average is slightly better than the blue 'treatment', indicating the DEX gain is your better option here.  

# Creating Monster Encounters
To help a DM compose a suitable encounter for a party, we started with a representative
set of Monster Manual XP values for various challenge-ratings, e.g. 
```{r, echo=FALSE, results='asis'}
knitr::kable(head(monsterCRExp, 10))
```
and the DM's-Guide table of multi-opponent XP-multipliers, e.g. 
```{r, echo=FALSE, results='asis'}
knitr::kable(head(monsterMultiCR, 10))
```
Given an XP budget set by difficulty-category encounter and party-composition, we can 
enumerate the various encounter sizes by CR-level, via the 

`targetXp <- partyEncounter( cts=c(3,1), lvls=c(5,7), difficulty='Hard' )
 listEncounters( targetXp )`

Which yields a table of how many monsters (up to 15) at various CRs come close to 
the target XP value.

## Multi-Challenge-Rating Encounters
To create a "boss-battle" with a few high-CR opponents and several minions, one approach
might be to split the tarketXp (e.g. 70%/30%) and call `listEncounters` twice, 
once for each sub-target value.  

(We don't yet support the variant that all minions have 1HP but otherwise normal AC, attack, and damage stats, but perhaps a DM could approximate this by taking a minion-count as if the CR were half the nominal value).  



You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.


Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
